
# Example 1: Simple linear regression (SLR)

The admissions committee of a comprehensive state university selected at random the records of 200 second-semester freshmen. The results, first-semester college GPA and SAT scores, are stored in the data set `Grades` in the package `PASWR`. The admissions committee wants to study the relationship between first-semester college grade point average (`gpa`) and scholastic aptitude test (`sat`) scores.

To open this data file, type:
```{r eval=FALSE}
install.packages("PASWR")
library(PASWR)
GRADES <- Grades
```

```{r echo=FALSE}
library(PASWR)
GRADES <- Grades
```

This will assign the data to the object `GRADES`. To view the data, type 

```{r eval=FALSE}
GRADES
```

into the script file and running that line by either placing the cursor on the line and pressing *Ctrl* and *Enter*, or by clicking on the *Run* button in the top-right corner of the script file. The data will then appear in the *Console* window below.

A spreadsheet of the data can be viewed by typing: 

```{r eval=FALSE}
fix(GRADES)
```

but note that you will need to close this window before trying to excute any further commands. The spreadsheet can also be viewed by clicking on the `GRADES` object in the Workspace (top right of the screen).

## Creating scatterplots

The relationship between `gpa` and `sat` scores can be examined by using a scatterplot:

```{r, eval=FALSE}
plot(gpa ~ sat, data = GRADES)
```

This will produce the basic scatterplot of `gpa` (y-axis) against `sat` (x-axis) shown in Figure \@ref(fig:grades-plot1).

```{r grades-plot1, echo = FALSE, fig.align='center', out.width="70%", fig.cap="Scatterplot of `gpa` versus `sat`."}
plot(gpa ~ sat, data = GRADES)
```

There are several options for annotating plots. First of all, you should add sensible axis labels to the plot. For example, by typing:

```{r, eval=FALSE}
plot(gpa ~ sat, data = GRADES, xlab = "SAT score", ylab = "GPA")
```

This will produce the scatterplot displayed in Figure \@ref(fig:grades-plot2).

```{r grades-plot2, echo = FALSE, fig.align='center', out.width="70%", fig.cap="Scatterplot of `gpa` versus `sat` with annotated axis labels."}
plot(gpa ~ sat, data = GRADES, xlab = "SAT score", ylab = "GPA")
```

Now edit the points by changing the symbol type, size and colour, such as

```{r, eval=FALSE}
plot(gpa ~ sat, data = GRADES, xlab = "SAT score", ylab = "GPA",
     pch = 2, cex = 1.5, col = "blue")
```

The plotting character is determined by the `pch` argument, the size of the plotted characters by `cex`, and the colour by the 1col` argument, to produce Figure \@ref(fig:grades-plot3).

```{r grades-plot3, echo = FALSE, fig.align='center', out.width="70%", fig.cap="Scatterplot of `gpa` versus `sat` with annotated axis labels and different characters plotted."}
plot(gpa ~ sat, data = GRADES, xlab = "SAT score", ylab = "GPA",
     pch = 2, cex = 0.7, col = "blue")
```

**DISCUSSION**: Looking at your plot, what can we say about the relationship between `gpa` and `sat`?

`r hide("Hint")` 
* Is the relationship linear or non-linear?
* Is the relationship positive or negative?
* How strong is the relationship?
`r unhide()`

## Fitting a linear model

To find the equation that best describes the relationship between `gpa` and `sat` use:
```{r eval=FALSE}
lm(gpa ~ sat, data = GRADES)
```

This fits a simple linear regression model with the response variable `gpa` and the explanatory variable `sat`.

From the `R` output, note down the equation of the fitted line that is given: *(Enter your answers to 3 decimal places.)*
<center> `gpa` = `r fitb(c("-1.19206381","-1.192064","-1.19206","-1.1921","-1.192","-1.19","-1.2"))` + `r fitb(c("0.003094","0.00309","0.0031","0.003"))` `sat` </center>

```{r eval = TRUE, webex.hide="Solution"}
model.lm<- lm(gpa ~ sat, data = GRADES)
coef(model.lm)
```

This is the line of best fit, describing the effect of `sat` on `gpa`.

<br>

**QUESTION**: Which of the following is the correct interpretation of the regression model?
```{r, echo = FALSE}
opts_p <- c("The intercept of the fitted line is -1.192, so for every unit increase in `sat`, we expect `gpa` to decrease by 1.192.", 
"The intercept of the fitted line is -1.192, so for every unit increase in `gpa`, we expect `sat` to decrease by 1.192.",
answer = "The slope coefficient of the fitted line is 0.0031, so for every unit increase in `sat`, we expect `gpa` to increase by 0.0031.",
"The coefficient of the fitted line is 0.0031, so for every unit increase in `gpa`, we expect `sat` to increase by 0.0031.")
```
`r longmcq(opts_p)`

<br>

**QUESTION**: Use `R` to calculate the point estimate of the change in the mean GPA when the SAT score increases by 50 points.
```{r eval = TRUE, webex.hide="Solution"}
coef(model.lm)[1]*50
```

<br>

**Plot of the data including the fitted line**

To obtain a plot of the fitted values, we need to store the regression in an object:

```{r}
model <- lm(gpa ~ sat, data = GRADES)
```

A plot of the data can be re-produced as before with the fitted line added using the `abline` command. This command uses the intercept and slope information from the fitted line in `model`. This is done using:

```{r eval=FALSE}
plot(gpa ~ sat, data = GRADES, xlab = "SAT score", ylab = "GPA")
abline(model)
```
